Notes on \\ Projected Knowledge Compilation \\ \today

### 2023-08-11 doc/pkc-notes.tex
Update optimization info

\item[Literals:]
  $\traverse(\lit, \phi) = \lit$ if $\dependencyset(\lit) \subseteq \dvarset$, and $\traverse(\lit, \phi) = \tautology$ if $\dependencyset(\lit) \in \pvarset$.
\item[Product Operations:]
  Let the operation be of the form $P(\lit_1, \lit_2, \ldots, \lit_k, \psi_1, \psi_2, \ldots, \psi_m)$, where each $\psi_i$ is either a sum or a  product operation.
  \begin{enumerate}
  \item Let $L = \{\lit_1, \lit_2, \ldots, \lit_k\}$.
  \item Define clausal formula $\phi'$ to be $\bcp(\simplify{\phi}{L})$
  \item For $1 \leq i \leq m$, recursively compute $\psi'_i = \traverse(\psi_i, \partition(\phi', \dependencyset(\psi_i)))$.  (Partitioning is only required when $m > 1$.)
  \item Return the product (with simplifications) for the set of arguments $\{\lit \in L \mid  \dependencyset(\lit) \subseteq \dvarset\} \cup \{ \psi'_i \mid  1 \leq i \leq m \}$.
  \end{enumerate}
\item[Sum Operations:]
  Let the operation be of the form $S(\lit_1, \lit_2)$ and having decision variable $x$.
  \begin{enumerate}
  \item For $1 \leq i \leq 2$, recursively compute $\psi'_i = \traverse(\psi_i, \phi)$.
  \item If $x \in \dvarset$, then return the sum (with simplifications) $S(\psi'_1, \psi'_2)$.
  \item Otherwise, let $\psi_{\forall} = \pkc(\ureduce{\phi}{x})$
    \begin{enumerate}
    \item If $\psi_{\forall} = \nil$, then return the sum (with simplifications) $S(\psi_1, \psi_2)$
    \item Otherwise return nested formula $S(\neg S(\neg \psi'_1, \psi_{\forall}), \psi'_2)$.
    \end{enumerate}
  \end{enumerate}
\end{description}

@@

We can improve the efficiency of the traversal by applying the following optimizations:
\begin{itemize}
\item Only data variables in $\psi$.  If $\dependencyset(\psi) \subseteq \dvarset$, then $\traverse(\psi, \phi) = \psi$.
\item Only data variables in $\phi$.  If $\dependencyset(\phi) \subseteq \dvarset$, then $\traverse(\psi, \phi) = \kc(\phi)$.
\item Only projection variables in $\phi$.  If $\dependencyset(\phi) \subseteq \pvarset$, then $\traverse(\psi, \phi) = \tautology$ if formula $\phi$ is satisfiable and $\nil$ otherwise.
\item Cached result. If $\psi$ is a sum operation and some previous traversal had the same clausal formula $\phi$ as its second argument, then the previously computed result can be reused.\footnote{
Caching for product operations is not, in general valid, since the two recursive calls made when processing a sum operation can be for product operations with the same clausal argument.}
\item Pure literal reduction.
Literal $\lit$ is said to be \emph{pure} in $\phi$ if some clause in $\phi$ contains $\lit$, and no clause in $\phi$ contains  $\obar{\lit}$.
Let $\pure(\phi)$ denote the set of all pure literals in $\phi$, and $\pure_{\pvarset}(\phi) = \{ \lit \in \pure(\phi) \mid  \dependencyset(\lit) \subseteq \pvarset \}$.
If $\pure_{\pvarset}(\phi) \not = \emptyset$, then $\phi$ can be reduced as
 $\phi \leftarrow \cfilter(\phi, \pure_{\pvarset})$.  Eliminating clauses can cause other literals to become pure, and so this reduction should be iterated until no pure literals remain.
\end{itemize}

The first two termination rules have some subtle differences.  For the
unoptimized version of $\traverse$, the arguments will always satisfy
$\dependencyset(\psi) \subseteq \dependencyset(\phi)$.  However, 
pure literal reduction can eliminate some literals from $\phi$
that remain in $\psi$, and so the two sets of variables can become
incomparable.  When both rules are applicable, it is more efficient to use the first, since it avoids an invocation of $\kc$.

>>>>

\item[Literals:]
  $\traverse(\lit, \phi) = \lit$ if $\dependencyset(\lit) \subseteq \dvarset$, and $\traverse(\lit, \phi) = \tautology$ if $\dependencyset(\lit) \in \pvarset$.
\item[Product Operations:]
  Let the operation be of the form $P(\lit_1, \lit_2, \ldots, \lit_k, \psi_1, \psi_2, \ldots, \psi_m)$, where each $\lit_i$ is a literal,
  and each $\psi_i$ is either a sum or a  product operation.
  \begin{enumerate}
  \item Let $L = \{\lit_1, \lit_2, \ldots, \lit_k\}$.
  \item Define clausal formula $\phi'$ to be $\bcp(\simplify{\phi}{L})$
  \item For $1 \leq i \leq m$, recursively compute $\psi'_i = \traverse(\psi_i, \partition(\phi', \dependencyset(\psi_i)))$.  (Partitioning is only required when $m > 1$.)
  \item Return the product (with simplifications) for the set of arguments $\{\lit \in L \mid  \dependencyset(\lit) \subseteq \dvarset\} \cup \{ \psi'_i \mid  1 \leq i \leq m \}$.
  \end{enumerate}
\item[Sum Operations:]
  Let the operation be of the form $S(\lit_1, \lit_2)$ and having decision variable $x$.
  \begin{enumerate}
  \item For $1 \leq i \leq 2$, recursively compute $\psi'_i = \traverse(\psi_i, \phi)$.
  \item If $x \in \dvarset$, then return the sum (with simplifications) $S(\psi'_1, \psi'_2)$.
  \item Otherwise, let $\psi_{\forall} = \pkc(\ureduce{\phi}{x})$
    \begin{enumerate}
    \item If $\psi_{\forall} = \nil$, then return the sum (with simplifications) $S(\psi_1, \psi_2)$
    \item Otherwise return nested formula $S(\neg S(\neg \psi'_1, \psi_{\forall}), \psi'_2)$.
    \end{enumerate}
  \end{enumerate}
\end{description}

@@

We can improve the efficiency of the traversal by applying the following optimizations:
\begin{itemize}
\item Only data variables in $\psi$.  If $\dependencyset(\psi) \subseteq \dvarset$, then $\traverse(\psi, \phi) = \psi$.
%%\item Only data variables in $\phi$.  If $\dependencyset(\phi) \subseteq \dvarset$, then $\traverse(\psi, \phi) = \kc(\phi)$.
\item Only projection variables in $\psi$.  If $\dependencyset(\psi) \subseteq \pvarset$, then $\traverse(\psi, \phi) = \tautology$ if formula $\phi$ is satisfiable and $\nil$ otherwise.
\item Cached result. If $\psi$ was an argument to a previous call to $\traverse$, then the previously computed result can be reused.
%%\item Pure literal reduction.
%%Literal $\lit$ is said to be \emph{pure} in $\phi$ if some clause in $\phi$ contains $\lit$, and no clause in $\phi$ contains  $\obar{\lit}$.
%%Let $\pure(\phi)$ denote the set of all pure literals in $\phi$, and $\pure_{\pvarset}(\phi) = \{ \lit \in \pure(\phi) \mid  \dependencyset(\lit) \subsete%%q \pvarset \}$.
%%If $\pure_{\pvarset}(\phi) \not = \emptyset$, then $\phi$ can be reduced as
%% $\phi \leftarrow \cfilter(\phi, \pure_{\pvarset})$.  Eliminating clauses can cause other literals to become pure, and so this reduction should be iterate%%d until no pure literals remain.
\end{itemize}

%% The first two termination rules have some subtle differences.  For the
%% unoptimized version of $\traverse$, the arguments will always satisfy
%% $\dependencyset(\psi) \subseteq \dependencyset(\phi)$.  However, 
%% pure literal reduction can eliminate some literals from $\phi$
%% that remain in $\psi$, and so the two sets of variables can become
%% incomparable.  When both rules are applicable, it is more efficient to use the first, since it avoids an invocation of $\kc$.
